---
title: 'Estimate migratory connectivity strength with the ''MigConnectivity'' package'
author: "Jeffrey A. Hostetler, Michael T. Hallworth"
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette: default
  rmarkdown::pdf_document: default
vignette: |
  %\VignetteIndexEntry{Estimate migratory connectivity strength with the ''MigConnectivity'' package}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: mc_library.bib 
csl: ecology.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval = FALSE}
install.packages('devtools')

devtools::install_github("SMBC-NZP/MigConnectivity", build_vignettes = TRUE)

library(MigConnectivity)
```

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(MigConnectivity)
library(sf)
set.seed(123)
```

## `estStrength` - Estimate the strength of migratory connectivity 

The <code>estStrength</code> function estimates the strength of migratory connectivity (MC) between populations during two different time periods within the annual cycle [@cohen_quantifying_2018; @roberts_migratory_2023]. Below, we illustrate how to estimate MC between three breeding and non-breeding regions. 

To estimate the strength of MC, you will need:   

1. to define the breeding and non-breeding regions  
1. the distance between the breeding regions and the distance between the non-breeding regions  
1. the transition probabilities between each breeding and non-breeding region (`psi`)  
1. the relative abundance within each of the regions where individuals originate from (here the breeding regions)
1. (optional) the total sample size of animals used to estimate the transition probabilities (provided automatically if psi comes from estTransition)


### Example 1. Estimate MC between breeding and non-breeding regions with both abundance and transition uncertainty

Simple example with four breeding and four nonbreeding regions and simulated data

Define the number of breeding and nonbreeding populations

```{r}
nBreeding <- 4 #number of breeding regions
nNonBreeding <- 4 #number of non-breeding regions
```

Generate distance matrices between the different regions
```{r}
originPos <- matrix(c(seq(-99, -81, 6),
                    rep(35, nBreeding)), nBreeding, 2)
targetPos <- matrix(c(seq(-93, -87, 2),
                    rep(9, nNonBreeding)), nNonBreeding, 2)

#distances between centroids of four breeding regions
breedDist <- distFromPos(originPos, surface = "ellipsoid")

#distances between centroids of four nonbreeding regions
nonBreedDist <- distFromPos(targetPos, surface = "ellipsoid")
```

```{r echo=FALSE}
par(bty="n")
plot(originPos[,1],originPos[,2],
     ylim=c(8.5,36),
     xlim=c(-100, -80),
     pch=19,
     axes=FALSE,
     yaxt="n",
     xaxt="n",
     cex=2,
     ylab="",
     xlab="Nonbreeding",
     main="Breeding")
points(targetPos[,1],targetPos[,2],pch=15,cex=2)
```  

Define the transition probabilities between the breeding and nonbreeding regions
and load previously estimated transition probabilities (using the package RMark)
```{r}
#transition probabilities form each breeding to each nonbreeding region
psiTrue <- samplePsis[["Low"]]

psiEst <- getCMRexample(1)

# Define total sample size for psi data 
# for small sample corrected version of MC 

n <- length(grep('[2-5]', psiEst$data$data$ch))

```
```{r echo=FALSE}
par(bty="n")
plot(originPos[,1],originPos[,2],
     ylim=c(8.5,36),
     xlim=c(-100, -80),
     pch=19,
     axes=FALSE,
     yaxt="n",
     xaxt="n",
     cex=2,
     ylab="",
     xlab="Nonbreeding",
     main="Breeding")
points(targetPos[,1],targetPos[,2],pch=15,cex=2)

for (b in 1:nBreeding) {
  for (nb in 1:nNonBreeding) {
    segments(originPos[b, 1], originPos[b, 2], 
             targetPos[nb, 1], targetPos[nb, 2],
             lwd = psiTrue[b, nb]*10, lty = b)
  }
}
```  

Define the relative abundance within the four breeding regions, and provide 
previously simulated and estimated relative abundances (simple BBS-style analyses 
using the functions `simCountData` and `modelCountDataJAGS`)
```{r}
#equal relative abundance among the four breeding regions, must sum to 1                      
relNTrue <- rep(1/nBreeding, nBreeding) 

relNEst <- abundExamples[[1]]

```
Calculate the true strength of migratory connectivity using the inputs above
```{r}
# Calculate the strength of migratory connectivity 
MCtrue <- calcMC(originDist = breedDist,
                 targetDist = nonBreedDist,
                 originRelAbund = relNTrue,
                 psi = psiTrue)

```

```{r}
MCtrue
```
Estimating the strength of MC by explicitly defining the total sample size of animals to estimate the transition probabilities

```{r}
MCest <- estStrength(originDist = breedDist,
             targetDist = nonBreedDist,
             originRelAbund = relNEst, 
             psi = psiEst,
             sampleSize = n,
             originSites = 5:8, 
             targetSites = c(3,2,1,4),
             nSamples = 1000)
```
```{r}
MCest
MCest$MC$mean - MCtrue
```

## Example 2. Estimate the strength of migratory connectivity after running estTransition 

```{r}
# Read in the processed Yellow Warbler data 
# see the Worked Examples Vignette for details on how data structured/created
newDir <- tempdir()
baseURL <- 'https://github.com/SMBC-NZP/MigConnectivity/blob/devpsi2/data-raw/YEWA/'
file.name <- "yewa_estTrans_data.rds"
url1 <- paste0(baseURL, file.name, '?raw=true')
temp <- paste(newDir, file.name, sep = '/')
utils::download.file(url1, temp, mode = 'wb')
YEWA_target_sites <- readRDS(temp)
YEWAdata <- readRDS(temp)
unlink(temp)

# take a quick look at the data # 
str(YEWAdata,1)
```

First, we estimate (psi) or transition probabilities using the `estTransition` function.

Note - in the Yellow Warbler data there are no overlapping data sources for individuals, i.e., we only had 
one type of data for each bird - so we set the overlap setting to “none”.

```{r, eval = FALSE}
## Run analysis for psi
psiYEWA <- estTransition(originSites = YEWAdata$originSites,
                         targetSites = YEWAdata$targetSites,
                         originPoints = YEWAdata$originPoints,
                         targetPoints = YEWAdata$targetPoints,
                         originAssignment = YEWAdata$originAssignment,
                         originNames = YEWAdata$originNames,
                         targetNames = YEWAdata$targetNames,
                         nSamples = 10,  # Set low for illustration
                         isGL = YEWAdata$isGL,
                         isTelemetry = YEWAdata$isTelemetry,
                         isRaster = YEWAdata$isRaster,
                         isProb = YEWAdata$isProb,
                         captured = YEWAdata$captured,
                         geoBias = YEWAdata$geoBias,
                         geoVCov = YEWAdata$geoVCov,
                         originRaster = YEWAdata$originRaster,
                         verbose = 0, 
                         maxTries = 400,
                         resampleProjection = YEWAdata$resampleProjection,
                         nSim = 30, 
                         dataOverlapSetting = "none",
                         targetRelAbund = YEWAdata$targetRelAbund,
                         returnAllInput = FALSE)
```

We then use the resulting object to estimate the strength of migratory connectivity. We first need some additional data, the distances between originSites and targetSites. 

```{r}
# calculate distance between originSites
originCentersYEWA <- st_centroid(YEWAdata$originSites)
originCentersYEWA <- st_transform(originCentersYEWA, 4326)
originDistYEWA <- distFromPos(st_coordinates(originCentersYEWA$geometry))

# calculate distance between targetSites 
targetCentersYEWA <- st_centroid(YEWAdata$targetSites)
targetCentersYEWA <- st_transform(targetCentersYEWA, 4326)
targetDistYEWA <- distFromPos(st_coordinates(targetCentersYEWA$geometry))
```

### Estimate the strength of Migratory Connectivity 

```{r, echo = FALSE}
newDir <- tempdir()
baseURL <- 'https://github.com/SMBC-NZP/MigConnectivity/blob/devpsi2/data-raw/YEWA/'
file.name <- "psiYEWA.rds"
url1 <- paste0(baseURL, file.name, '?raw=true')
temp <- paste(newDir, file.name, sep = '/')
utils::download.file(url1, temp, mode = 'wb')
psiYEWA <- readRDS(temp)
unlink(temp)

```


```{r}
# Estimate the strength of migratory connectivity
YEWAmc <- estStrength(originDist = originDistYEWA,
                      targetDist = targetDistYEWA,
                      originRelAbund = YEWAdata$originRelAbund,
                      psi = psiYEWA,
                      nSamples = 5000,
                      verbose = 0)
YEWAmc
```

```{r echo = FALSE}
#saveRDS(YEWAmc, "YEWAmc.rds")
```

### Literature Cited

<div id="refs"></div>

