---
title: '''MigConnectivity'' package: isotope simulation'
author: "Michael T. Hallworth, Jeffrey A. Hostetler"
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette: default
  rmarkdown::pdf_document: default
vignette: |
  %\usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{MigConnectivity package: isotope simulation}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = "#>", 
                      tidy = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

Load required packages
```{r eval = FALSE}
library(raster)
library(MigConnectivity)
library(rgeos)
```

The following is a simulation that tests the how the spatial arrangement of target sites influences MC from stable-hydrogen isotopes. The following simulation is run using data generated within the code but we use the Ovenbird as an example species. 

### Ovenbird distribution
Read in the Ovenbird distribution and create a species distribution map from the abundance data. 
```{r eval = FALSE}
# read in raster layer
download.file(paste0("https://raw.githubusercontent.com/SMBC-NZP/MigConnectivity",
                     "/master/data-raw/Spatial_Layers/bbsoven.txt"),
              destfile = "bbsoven.txt")

OVENabund <- raster::raster("bbsoven.txt")

OVENdist <- OVENabund
OVENdist[OVENdist>0]<-1
OVENdist[OVENdist==0]<-NA

OVEN_single_poly <- raster::rasterToPolygons(OVENdist, dissolve = TRUE)
raster::crs(OVEN_single_poly) <- MigConnectivity::projections$WGS84
raster::crs(OVENabund) <- MigConnectivity::projections$WGS84
```
To complete the simulation we need a template to ensure the raster resolution is the same as the assignment raster. To do this, we use the isotope data as our template. We grab the isotope base-map using the <code>getIsoMap</code> function. 
```{r eval = FALSE}
### get isotope raster for template
rasterTemplate <- MigConnectivity::getIsoMap()
```

### Crop template to distribution and mask with distribution
```{r eval = FALSE}
rasterTemplate <- raster::mask(raster::crop(rasterTemplate, OVEN_single_poly),
                               OVEN_single_poly)

# rasterize the distribution for relative abundance so that raster
# dimensions and resolution match the isotope layer
relativeAbund <- projectRaster(OVENabund,rasterTemplate)
relativeAbund  <- relativeAbund/cellStats(relativeAbund,sum)
```

## Generate target sites
The simulation is focused on the effect of target sites on MC when using stable-hydrogen isotopes. The following code generates various target site layers used in the simulation.
```{r eval = FALSE}
# generate target sites
targetRanges <- vector('list',5)
# 3' latitude
targetRanges[[1]] <- rasterToPolygons(raster(xmn = -180,
                                         xmx = -40,
                                         ymn = 25,
                                         ymx = 85,
                                         res = c(140,3)))

# 5' latitude
targetRanges[[2]] <- rasterToPolygons(raster(xmn = -180,
                                             xmx = -40,
                                             ymn = 25,
                                             ymx = 85,
                                             res = c(140,5)))

# 10' latitude
targetRanges[[3]] <- rasterToPolygons(raster(xmn = -180,
                                         xmx = -40,
                                         ymn = 25,
                                         ymx = 85,
                                         res = c(140,10)))

# 12 isotope units
featherIso <- (0.95*getIsoMap(period = "GrowingSeason")+(-17.57))
iso <- crop(featherIso, extent(c(-180,-40,25,85)))
isocut <- cut(iso, breaks= seq(cellStats(iso,min),cellStats(iso,max),12))
targetRanges[[4]] <- rasterToPolygons(isocut, dissolve = TRUE)


# 12*2 isotope units
isocut <- cut(iso, breaks= seq(cellStats(iso,min),cellStats(iso,max),24))
targetRanges[[5]] <- rasterToPolygons(isocut, dissolve = TRUE)



# Keep only the targetSites that intersect with the OVEN polygon
targetRanges <- lapply(targetRanges,raster::intersect,OVEN_single_poly)
targetRanges <- lapply(targetRanges, sp::spTransform,CRS(projections$WGS84))

for(i in 1:5){
targetRanges[[i]]$Target <- 1:nrow(targetRanges[[i]])
}
```

#### Generate simulated data
```{r, eval = FALSE}
#Generate random breeding locations using the 10' target sites
Site1 <- spsample(targetRanges[[3]][1:2,],n=100,type = "random")
Site2 <- spsample(targetRanges[[3]][2:3,],n=100, type = "random")
Site3 <- spsample(targetRanges[[3]][2:4,], n= 100, type = "random")

# Capture coordinates
capCoords <- array(NA,c(3,2))
capCoords[1,] <- cbind(-98.17,28.76)
capCoords[2,] <- cbind(-93.70,29.77)
capCoords[3,] <- cbind(-85.000,29.836)

featherIso <- (0.95*getIsoMap(period = "GrowingSeason")+(-17.57))

# Extract simulated data
iso_dat <- data.frame(Site = rep(1:3,each = 100),
                      xcoords = c(Site1@coords[,1],Site2@coords[,1],Site3@coords[,1]),
                      ycoords = c(Site1@coords[,2],Site2@coords[,2],Site3@coords[,2]),
                      targetSite = over(rbind(Site1,Site2,Site3),
                                        targetRanges[[3]])$Target,
                      featherIso = extract(featherIso,rbind(Site1,Site2,Site3)))

iso_dat <- iso_dat[complete.cases(iso_dat),]

# generate transition data from simulation
sim_psi <- table(iso_dat$Site,iso_dat$targetSite)

for(i in 1:nrow(sim_psi)){
  sim_psi[i,]<-sim_psi[i,]/sum(sim_psi[i,])
}

states <- raster::getData("GADM", country = "United States", level = 1)
originSites <- states[(states$NAME_1 %in% c("Texas","Louisiana","Florida")),]
originSites <- sp::spTransform(originSites, 
                               sp::CRS(MigConnectivity::projections$WGS84))

originDist <- MigConnectivity::distFromPos(rgeos::gCentroid(originSites,
                                                            byid = TRUE)@coords)
targetDistMC <- MigConnectivity::distFromPos(rgeos::gCentroid(targetRanges[[3]], 
                                                              byid = TRUE)@coords)
```

## Start of simulations

*warning* this takes a long time to run. 

```{r, eval = FALSE}
originRelAbund <- rep(1/3, 3)
nTargetSetups <- 5
nSims <- 1            #SET LOW FOR EXAMPLE
# nSims <- 200        
nOriginSites = 3

targetPoints0 <- matrix(NA, 300, 2)

sim.output <- data.frame(targetSetup = rep(NA,nSims),
                         sim = rep(NA,nSims),
                         MC.generated = rep(NA,nSims),
                         MC.realized = rep(NA,nSims),
                         MC.est = rep(NA,nSims),
                         MC.low = rep(NA,nSims),
                         MC.high = rep(NA,nSims),
                         rM.realized = rep(NA,nSims),
                         rM.est = rep(NA,nSims),
                         rM.low = rep(NA,nSims),
                         rM.high = rep(NA,nSims))


library(doSNOW);library(foreach)
cl <- parallel::makeCluster(5)
doSNOW::registerDoSNOW(cl)


Sys.time()
a <- Sys.time()
output.sims <- foreach(target=1:nTargetSetups,
                       .combine = rbind,
                       .export = ls(.GlobalEnv),
                       .packages = c("raster","sp","MigConnectivity")) %dopar% {
#for(target in 1:nTargetSetups){
  set.seed(9001)
  targetSites <- targetRanges[[target]]
  targetDist <- distFromPos(rgeos::gCentroid(targetSites, byid = TRUE)@coords)
  # Extract simulated data
  iso_dat <- data.frame(Site = rep(1:3,each = 100),
                        xcoords = c(Site1@coords[,1],Site2@coords[,1],Site3@coords[,1]),
                        ycoords = c(Site1@coords[,2],Site2@coords[,2],Site3@coords[,2]),
                        targetSite = over(rbind(Site1,Site2,Site3),targetSites)$Target,
                        featherIso = extract(featherIso,rbind(Site1,Site2,Site3)))

  iso_dat <- iso_dat[complete.cases(iso_dat),]

  # generate transition data from simulation
  sim_psi <- prop.table(table(iso_dat$Site,factor(iso_dat$targetSite,
                                                  1:nrow(targetSites))), 1)

  MC.generated <- calcMC(originDist = originDist,
                          targetDist = targetDist,
                          originRelAbund = originRelAbund,
                          psi = sim_psi)

  for (sim in 1:nSims) {
    cat("Simulation Run", sim, "of", nSims, "for target",target,"at", date(), "\n")
    sim_move <- simMove(rep(100, nOriginSites), originDist, targetDist, sim_psi, 1, 1)
    originAssignment <- sim_move$animalLoc[,1,1,1]
    targetAssignment <- sim_move$animalLoc[,2,1,1]
    for (i in 1:300) {
      targetPoint1 <- spsample(targetSites[targetAssignment[i],], 
                               n= 1, type = "random", iter = 25)
      targetPoints0[i,] <- targetPoint1@coords
    }
    targetPoints <- SpatialPoints(targetPoints0, CRS(projections$WGS84))
  
    # Extract simulated data
    iso_dat <- data.frame(Site = originAssignment,
                          xcoords = targetPoints0[,1],
                          ycoords = targetPoints0[,2],
                          targetSite = over(targetPoints,targetSites)$Target,
                          featherIso = extract(featherIso,targetPoints))
  
    iso_dat <- iso_dat[complete.cases(iso_dat),]
  
    # generate transition data from simulation
    sim_psi0 <- table(iso_dat$Site,
                      factor(iso_dat$targetSite,
                             min(targetSites$Target):max(targetSites$Target)))
    sim_psi_realized <- prop.table(sim_psi0, 1)
  
    # get points ready for analysis
    nSite1 <- table(iso_dat$Site)[1]
    nSite2 <- table(iso_dat$Site)[2]
    nSite3 <- table(iso_dat$Site)[3]
  
    nTotal <- nSite1+nSite2+nSite3
  
    originCap <- array(NA, c(nTotal,2))
  
    wherecaught <- rep(paste0("Site", 1:3), c(nSite1, nSite2, nSite3))
  
    originCap[which(pmatch(wherecaught,"Site1",dup = TRUE)==1),1] <- capCoords[1,1]
    originCap[which(pmatch(wherecaught,"Site1",dup = TRUE)==1),2] <- capCoords[1,2]
  
    originCap[which(pmatch(wherecaught,"Site2",dup = TRUE)==1),1] <- capCoords[2,1]
    originCap[which(pmatch(wherecaught,"Site2",dup = TRUE)==1),2] <- capCoords[2,2]
  
    originCap[which(pmatch(wherecaught,"Site3",dup = TRUE)==1),1] <- capCoords[3,1]
    originCap[which(pmatch(wherecaught,"Site3",dup = TRUE)==1),2] <- capCoords[3,2]
  
    originPoints <- SpatialPoints(originCap)
    crs(originPoints) <- sp::CRS(MigConnectivity::projections$WGS84)
  
    MC.realized <- calcMC(originDist = originDist,
                          targetDist = targetDist,
                          originRelAbund = originRelAbund,
                          psi = sim_psi_realized,
                          sampleSize=nTotal)
  
  
  
    originPointDists <- distFromPos(originCap)
    targetPointDists <- distFromPos(cbind(iso_dat$xcoords, iso_dat$ycoords))
  
  
    simAssign <- isoAssign(isovalues = iso_dat$featherIso,
                           isoSTD = 12,
                           intercept = -17.57,
                           slope = 0.95,
                           odds = 0.67,
                           restrict2Likely = FALSE,
                           nSamples = 500,
                           sppShapefile = OVEN_single_poly,
                           relAbund = relativeAbund,
                           verbose = 0,
                           isoWeight = -0.7,
                           abundWeight = 0,
                           assignExtent = c(-179,-60,15,89),
                           element = "Hydrogen",
                           surface = FALSE,
                           period = "GrowingSeason")
  
    simEst <- estMC(originRelAbund = originRelAbund,
                    targetDist = targetDist,
                    targetIntrinsic = simAssign,
                    targetSites = targetSites,
                    originPoints = originPoints,
                    originSites = originSites,
                    originDist = originDist,
                    nSamples = 100,
                    verbose = 0,
                    calcCorr = TRUE,
                    alpha = 0.05,
                    approxSigTest = FALSE,
                    sigConst = 0,
                    isIntrinsic = TRUE,
                    nSim = 5)
  
    sim.output$targetSetup[sim] <- target
    sim.output$sim[sim]<-sim
    sim.output$MC.generated[sim] <- MC.generated
    sim.output$MC.realized[sim] <- MC.realized
    sim.output$MC.est[sim] <- simEst$meanMC
    sim.output$MC.low[sim] <- simEst$bcCI[1]
    sim.output$MC.high[sim] <- simEst$bcCI[2]
    sim.output$rM.realized[sim] <- ncf::mantel.test(originPointDists, 
                                                    targetPointDists,
                                                    resamp=0, 
                                                    quiet = TRUE)$correlation
    sim.output$rM.est[sim] <- simEst$meanCorr
    sim.output$rM.low[sim] <- simEst$bcCICorr[1]
    sim.output$rM.high[sim] <- simEst$bcCICorr[2]
  
  }
  return(sim.output)
}
Sys.time()-a

stopCluster(cl)
```
Summarize the output 

```{r, eval = FALSE}
sim.output <- transform(output.sims,
                        MC.generated.cover = as.integer((MC.low <= MC.generated & 
                                                           MC.high >= MC.generated)),
                        MC.realized.cover = as.integer((MC.low <= MC.realized & 
                                                          MC.high >= MC.realized)),
                        MC.generated.error = MC.est - MC.generated, 
                        MC.realized.error = MC.est - MC.realized,
                        rM.cover = as.integer((rM.low <= rM.realized & 
                                                 rM.high >= rM.realized)),
                        rM.error = rM.est - rM.realized)

summary(sim.output)
# Examine results
aggregate(MC.generated.error ~ targetSetup, sim.output, mean)
aggregate(MC.generated.error ~ targetSetup, sim.output, function (x) mean(abs(x)))
aggregate(MC.generated.cover ~ targetSetup, sim.output, mean)
aggregate(MC.realized.error ~ targetSetup, sim.output, mean)
aggregate(MC.realized.error ~ targetSetup, sim.output, function (x) mean(abs(x)))
aggregate(MC.realized.cover ~ targetSetup, sim.output, mean)
aggregate(I(MC.realized - MC.generated) ~ targetSetup, sim.output, mean)
aggregate(rM.error ~ targetSetup, sim.output, mean)
aggregate(rM.error ~ targetSetup, sim.output, function (x) mean(abs(x)))
aggregate(rM.cover ~ targetSetup, sim.output, mean)
aggregate(MC.est ~ targetSetup, sim.output, mean)
```
