---
title: 'Preparing data to estimate transition probabilites and/or migratory connecitivity using the ''MigConnectivity'' package'
author: "Jeffrey A. Hostetler, Michael T. Hallworth"
date: '`r Sys.Date()`'
output:
  rmarkdown::html_vignette: default
  rmarkdown::pdf_document: default
vignette: |
  %\VignetteIndexEntry{MigConnectivity package}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: mc_library.bib 
csl: ecology.csl
---
```{r, echo = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
```

This vignette is intended to help users structure their data for analysis using the 
MigConnectivity package. There are many ways to derive the data needed for analysis and below, we provided examples and code as a reference for users. However, the most appropriate methods and approaches for particular questions and data types should be considered by 
the users. 

There are examples at the end of this vignette that illustrate ways to calculate 
bias and variance metrics using light-level geolocator data derived from the
[SGAT](#SGAT_bias) and [FLightR](#FLightR_bias) packages.


## Using data integration to estimate the strength of migratory connectivitiy 

This MigConnectivity package provides tools for estimating the pattern and strength
of migratory connectivity. We've developed methods to integrate intrinsic markers,
tracking, and long-distance band reencounter data collected from the same or
different individuals. The package functionality includes integrated analyses
that account for differences in precision, bias, reencounter probability, 
and directionality among different data types (banding reencounter, 
light-level geolocator, stable isotopes in tissues, genetics, and GPS). The package
separates estimation of migratory connectivity into two functions: estTransition and estStrength. 
The `estTransition` function estimates transition probabilities (ψ) from seasonal movement data (i.e. pattern of movement) 
and the `estStrength` estimates the strength of migratory connectivity (MC) from transition probabilities.
Below we provide a worked example using Yellow Warbler (<i>Setophaga petechia</i>) data to help users format their data to use the new 
functionality in the MigConnectivity package. There are multiple published Yellow Warbler data sets that can be used to estimate migratory connectivity. Here, we illustrate how to integrate stable-isotopes, genetic data, and light-level geolocators to estimate MC.

### Define regions

The Yellow Warbler genoscape identifies 5 genetically distinct breeding populations (Arctic, Pacific Northwest, Southwest, Central, and East) from 419 sampled individuals at 50 breeding locations [@bay_genetic_2021]. We identified 4 nonbreeding regions using ecoregions which include the Pacific and Central Mexico, Atlantic Lowland Mexico, Central America and South America. 

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap="Yellow Warbler regions. The breeding regions are defined by genoscape-defined populations and the nonbreeding regions were identified by ecoregions across the ranges"}

# THIS CHUNK IS NOT PRINTED TO THE VIGNETTE # 
# BUT READS IN SOME DATA NEEDED TO RUN THE ANALYSIS # 

library(MigConnectivity)
library(sf)
library(terra)

set.seed(1)
YEWA_target_sites <- readRDS("C:/Users/mhall/OneDrive/Desktop/YEWA data/yelwar_ecoregions/YEWA_target_sites.rds")

YEWA_target_sites <- YEWA_target_sites[,c("Region","targetSite","geometry")]

# vector of target names #
targetNames <- c("Pacific and Central Mexico", "Atlantic Lowland Mexico",
                 "Central America", "South America")



# Define the origin sites
YEWA_origin_sites <- sf::st_read("C:/Users/mhall/OneDrive/Desktop/YEWA data/yelwar_genoscape/originSitesYEWA.shp", 
                                 quiet = TRUE)
# vector of origin names 
originNames <- c("Arctic", "Pacific Northwest", "Southwest", "Central", "East")

# object with the number of originSites 
nOriginSites <- 5

#Plot function
lims <- rbind(st_bbox(YEWA_target_sites),st_bbox(YEWA_origin_sites))

mins <- apply(lims[,c(1,2)], 2, FUN = "min")
maxs <- apply(lims[,c(3,4)], 2, FUN = "max")

par(mar = c(0,0,0,0))
plot(YEWA_origin_sites$geometry, 
     ylim = c(mins[2],maxs[2]), 
     xlim = c(mins[1],maxs[1]), 
     col = viridis::viridis(n = 5, option = "G", alpha = 0.25), 
     border = "white", 
     lwd = 0.25)
plot(YEWA_target_sites$geometry, col = viridis::viridis(n = 4, option = "D"), border = "white", lwd = 0.25, add = TRUE)
```

### Define Origin Sites

The Yellow Warbler has published genoscape and isotope data on migratory movements from birds captured and sampled in the nonbreeding season and geolocators on birds captured in the breeding season. Genetic sampling of 203 individuals occurred at 20 locations across the nonbreeding range, from Baja Mexico, south through central Mexico, Central America (Costa Rica and Nicaragua), and South America (Trinidad and Venezuela) [@bay_genetic_2021] and the isotope samples were collected at one nonbreeding site in western Mexico (n = 210) [@valdez-juarez_use_2018]. 

Below we create a vector of the origin names. These names correspond with the genoscape. 
```{r, eval = FALSE}
# Load shapefile of origin sites derived from genoscape data
YEWA_origin_sites <- sf::st_read("originSitesYEWA.shp", 
                                 quiet = TRUE)
# vector of origin names 
originNames <- c("Arctic", "Pacific Northwest", "Southwest", "Central", "East")

# object with the number of originSites 
nOriginSites <- 5
```

```{r, echo = FALSE, fig.cap="Yellow Warbler regions. The breeding regions are defined by genoscape-defined populations and the nonbreeding regions were identified by ecoregions across the ranges"}

par(mar = c(0,0,0,0))
plot(YEWA_origin_sites$geometry, 
     ylim = c(mins[2],maxs[2]), 
     xlim = c(mins[1],maxs[1]), 
     col = viridis::viridis(n = 5, option = "G"), 
     border = "white", 
     lwd = 0.25)
plot(YEWA_target_sites$geometry, 
     col = viridis::viridis(n = 4, option = "D", alpha = 0.25),
     border = "white", lwd = 0.25, add = TRUE)
```

### Sample data 

For information about sampling and assignment methods see Bay et al. (2021) for genetics, Valdez-Juárez et al. (2018) for isotopes, and Witynski and Bonter (2018) for geolocators.

### Load the genetic sample data

```{r, eval = FALSE}

#read in the data 
genetic_samples <- read.delim("YEWA_nonbreeding_genetic_samples.txt",
                              stringsAsFactors = TRUE)

str(genetric_samples,1)
```

```{r, echo = FALSE}
genetic_samples <- read.delim("C:/Users/mhall/OneDrive/Desktop/YEWA data/YEWA_genetic_samples/YEWA.rep_indiv_est.Wintering.w_meta.txt",
                              stringsAsFactors = TRUE)

names(genetic_samples) <- gsub(names(genetic_samples), pattern = "X[0-9]..", replacement = "")

str(genetic_samples,1)
```

The genetic data have the probabilty that each individual is assigned to 
a specific genetic population. These data are used during the resampling
process to estimate transition probabilities and the strength of MC. Below, 
we format the data so we can easily use it in our analysis. 

```{r}
# Note that we re-order the probabilites to match the order of originNames #
YEWA_genetics <- as.matrix(genetic_samples[,c("Alaska","WesternBoreal",
                                              "Southwest","Central",
                                              "East")])

# give names to the object
dimnames(YEWA_genetics) <- list(as.character(genetic_samples$indiv), 
                                originNames)

# Double-check that all rows sum to 1
summary(rowSums(YEWA_genetics))
```

Convert the data into a spatial object 
```{r}
# Convert the data into spatial object 
targetPoints_genetic <- st_as_sf(genetic_samples[,c("longitude","latitude")],
                                  coords = c("longitude","latitude"),
                                  crs = 4326) # WGS84

# Assign the data type to the spatial layer 

targetPoints_genetic$Type <- "Genetics Capture"
```


### Load the hydrogen stable isotope data 

```{r, eval = FALSE}
isotope_samples <- read.csv("YEWA isotopes Jalisco.csv")

str(isotope_samples)
```

```{r, echo = FALSE}
isotope_samples <- read.csv("C:/Users/mhall/OneDrive/Desktop/YEWA data/YEWA_isotope/YEWA isotopes Jalisco.csv")

isotope_samples <- isotope_samples[,c("Delta","Longitude","Latitude")]
```

convert the data into a spatial object 

```{r}
# Convert the data into spatial object 
targetPoints_iso <- st_as_sf(isotope_samples[,c("Longitude","Latitude")],
                                  coords = c("Longitude","Latitude"),
                                  crs = 4326) # WGS84

# Assign the data type to the spatial layer 

targetPoints_iso$Type <- "Isotope Capture"

```


For hydrogen stable isotope, we generate probabilistic origin surfaces (i.e., rasters) for each isotope sample using the `isoAssign` function available in the R package MigConnectivity. 

```{r, eval = FALSE}
# Make isotope assignments 
isoRasterYEWA <- isoAssign(isovalues = isotope_samples$Delta,
                           isoSTD = 18.3,
                           intercept = -28.67,
                           slope = 0.59,
                           sppShapefile = YEWA_origin_sites)

```

```{r, echo = FALSE}
yewa_data <- readRDS("yewa_estTrans_data.rds")

isoRasterYEWA <- yewa_data$originRaster
```

### Load the light-level geolocator data 

We obtained location estimates from seven individuals tracked using light-level geolocators deployed and recovered in Maine (n = 4) and Wisconsin (n = 3; Witynski & Bonter, 2018). We used the daily location estimates while individuals remained on the breeding grounds to derive location bias and uncertainty estimates. The geolocator data were processed and reported by @witynski_crosswise_2018.

```{r}
# Capture locations #
# See dates and summary data from Table 1. page 44.
# Maine #129
# Maine #132
# Maine #136
# Maine #137
# Wisconsin #139
# Wisconsin #144
# Wisconsin #146

caplocs <- cbind(c(rep(-70.6142,4),rep(-87.8123,3)),
                 c(rep(42.9891,4),rep(42.5000,3)))

# Departures #
departs <- as.POSIXct(c("2015-09-04",
                        "2015-08-22",
                        "2015-09-04",
                        "2015-08-28",
                        "2015-08-24",
                        "2015-08-24",
                        "2015-08-30"))
```

```{r, echo = FALSE}
# read in the daily location data #
EstLocs <- read.csv("C:/Users/mhall/OneDrive/Desktop/YEWA data/yelwar_geolocators/Yellow Warbler_Witynski (1).csv")

# make it spatial 
YEWA_gl_locs <- sf::st_as_sf(EstLocs,
                             coords = c("location.long",
                                        "location.lat"),
                             crs = 4326)
```

```{r, eval = FALSE}
# read in the daily location data #
EstLocs <- read.csv("Yellow Warbler_Witynski.csv")

# make it spatial 
YEWA_gl_locs <- sf::st_as_sf(EstLocs,
                             coords = c("location.long",
                                        "location.lat"),
                             crs = 4326)
```


```{r}
# Summarized nonbreeding locations reported in Witynski and Bonter

YEWA_GL_nb <- sf::st_as_sf(data.frame(bird = factor(c(129,132,136,137,139,144,146)),
                                      long = c(-75.09,-74.81,-74.30,-74.21,-68.72,-67.08,-65.39),
                                      lat = c(10.03,8.52,10.51,3.14,0.96,5.73,6.67)),
                               coords = c("long","lat"),
                               crs = 4326)

# assign type to GL spatial object
YEWA_GL_nb$Type <- "GL Estimated Target"

# Convert the capture locations to a spatial object
YEWA_GL_breed <- st_as_sf(data.frame(bird = c(129,132,136,137,139,144,146),
                                  longitude = caplocs[,1],
                                  latitude = caplocs[,2]),
                       coords = c("longitude","latitude"),
                       crs = 4326)

# assign type to GL spatial object
YEWA_GL_breed$Type <- "GL Capture"

```

Calculate location bias and covariance which are used to estimate transition 
probabilities (`estTransition`). 

```{r}
# project from WGS84 into meters #
YEWAmeters <- sf::st_transform(YEWA_gl_locs,'ESRI:102010')
CapMeters <- sf::st_transform(YEWA_GL_breed, 'ESRI:102010')

# split birds into their own data #
YEWAests <- split(YEWAmeters,
                  f = YEWAmeters$'individual.local.identifier')

# Empirical estimates of bias and vcov, but on mean locations by bird
ests2 <- mapply(x = YEWAests,
                lats = st_coordinates(CapMeters)[,2],
                lng = st_coordinates(CapMeters)[,1],
                z = departs,
                FUN = function(x,lats,lng,z){
              # keep only data from deployment until departure from breeding #
                  newdata <- x[x$timestamp<z,]

              # calculate the difference in meters from estimated location
              # to actual capture location
                  long <- mean(st_coordinates(newdata)[,1]) - lng
                  lat <- mean(st_coordinates(newdata)[,2]) - lats

                  return(c(difflong = long,
                                    difflat = lat))},

               SIMPLIFY = TRUE)

(geoBiasYEWA <- apply(ests2, 1, mean)) #Bias

(geoVCovYEWA <- cov(t(ests2))) #Variance-covariance matrix

```

### Set up input to run estTransition 

The following code creates objects that are needed and/or helpful when using 
the `estStrength` or `estTransition` functions. 

```{r}
## Set up input for running in estTransition

# Number of geolocator birds 
nGL <- nrow(YEWA_GL_breed)

# We have 0 birds with telemetry data
nTelemetry <- 0

# How many birds have raster data (isotopes)
nRaster <- nrow(isotope_samples)

# How many birds have probability data (genetics)
nProb <- nrow(YEWA_genetics)

# Total number of animals 
nAnimals <- nGL + nTelemetry + nRaster + nProb

# Set up a vector of TRUE/FALSE indicating if the individual 
# has geolocator data 
isGLYEWA <- rep(c(TRUE, FALSE), c(nGL, nTelemetry + nRaster + nProb))

# Vector indicating whether an individual has telemetry data 
isTelemetryYEWA <- rep(FALSE, nAnimals)

# Vector indicating whether an individual has raster assignment data 
isRasterYEWA <- rep(c(FALSE, TRUE, FALSE), c(nGL + nTelemetry, nRaster, nProb))

# Vector indicating whether an individual has probability data 
isProbYEWA <- rep(c(FALSE, TRUE), c(nGL + nTelemetry + nRaster, nProb))
```

Now that we have objects that tell the function which type of data are associated
with each individual we can set up the spatial data. 

```{r}
# combine all the target locations into a single object 
# NOTE - we reorder the columns to match between the data sets
targetPointsYEWA <- rbind(YEWA_GL_nb[,c("Type","geometry")],
                          targetPoints_iso[,c("Type","geometry")],
                          targetPoints_genetic[,c("Type","geometry")])

# Transform (re-project the data) into Equidistant conic projection
targetPointsYEWA <- st_transform(targetPointsYEWA, "ESRI:102010")

# A vector indicating where the birds were captured 
capturedYEWA <- rep(c("origin", "target"), c(nGL, nTelemetry + nRaster + nProb))
```

Here we do some spatial analyses to determine target sites from the data we've 
gathered so far. These data are needed for the analysis. 

```{r}
# Where birds were captured in the origin sites (i.e., geolocators)
originPointsYEWA <- CapMeters

# Transform (re-project) data into Equidistant conic projection 
targetSitesYEWA <- st_transform(YEWA_target_sites, "ESRI:102010")

# Create a vector of the target assignments of the input data
targetAssignment <- suppressMessages(unclass(sf::st_intersects(x = targetPointsYEWA,
                                                               y = targetSitesYEWA,
                                                               sparse = TRUE)))

# Set any target assignment that results in 0 to NA
targetAssignment[lengths(targetAssignment)==0] <- NA

# convert from a list to an array #
targetAssignment <- array(unlist(targetAssignment))

# If the assignment is NA - assign to the nearest targetSite
targetAssignment[is.na(targetAssignment)] <- sf::st_nearest_feature(x = targetPointsYEWA[is.na(targetAssignment),],
                                                                    y = targetSitesYEWA)
```

### Abundance estimates 

Accounting for relative abundance is an important component when estimating 
migratory connectivity strength or estimating transition probabilities between 
two or more phases of the annual cycle. 

Below, we use eBird.org products to estimate abundance. Note - users may need 
to create an account and get permission to access eBird data. See information on
to [Download eBird data products](https://science.ebird.org/en/use-ebird-data/download-ebird-data-products).

```{r}
# Read in data from eBird 
library(ebirdst)
library(tidyverse)

# read in the eBird median abundance weekly cube data
rastAbundMed <- raster::brick("yelwar-ERD2019-STATUS-20200930-e26043a8_lr_2019_abundance_median.tif")

# assign the correct dates to the raster layers
rastAbundMed <- label_raster_stack(rastAbunMed)
```

```{r}
# dates for each abundance layer 
weeks <- parse_raster_dates(rastAbundMed)

run_review <- subset(ebirdst_runs, 
                     species_code == "yelwar")

yelwar_dates <- run_review %>%
  # just keep the seasonal definition columns
  select(setdiff(matches("(start)|(end)"), matches("year_round"))) %>%
  # transpose
  gather("label", "date") %>%
  # spread data so start and end dates are in separate columns
  separate(label, c("season", "start_end"), "_(?=s|e)") %>%
  spread(start_end, date) %>%
  select(season, start_dt = start, end_dt = end) %>%
  filter(season != "resident")
  
  # did the season pass review
quality_rating <- run_review[paste0(yelwar_dates$season, "_quality")]

yelwar_dates$pass <- as.integer(quality_rating) > 1

# Get the weeks that represent breeding and non-breeding
weeks_season <- rep(NA_character_, length(weeks))
for (i in seq_len(nrow(yelwar_dates))) {
  s <- yelwar_dates[i, ]
  # skip seasonal assignment if season failed
  if (!s$pass) {
    next()
  }
  # handle seasons cross jan 1 separately
  if (s$start_dt <= s$end_dt) {
    in_season <- weeks >= s$start_dt & weeks <= s$end_dt
  } else {
    in_season <- weeks >= s$start_dt | weeks <= s$end_dt
  }
  weeks_season[in_season] <- s$season
}

b <- which(weeks_season == 'breeding')
w <- which(weeks_season == 'nonbreeding')
```

```{r}

# Project the target and origin sites in the projection of the eBird abundance
# so abundance data aren't changed while projecting 
originSitesProj <- sf::st_transform(YEWA_origin_sites, st_crs(rastAbundMed))
targetSitesProj <- sf::st_transform(YEWA_target_sites, st_crs(rastAbundMed))
```


### Extract nonbreeding relative abundance estimates

```{r}
# extract winter raster values for each region 
system.time(
weekly_region_vals  <- raster::extract(rastAbundMed[[w]],targetSitesProj)
)

str(weekly_region_vals)

sumByWeek <- lapply(weekly_region_vals, 
                    FUN = function(x){apply(x, 2, sum, na.rm = TRUE)})

regionSum <- lapply(sumByWeek, max)

relAbundYEWAWint <- prop.table(unlist(regionSum))

```

Extract breeding relative abundance estimates

```{r}
# extract breeding raster values for each region 
system.time(
weekly_region_vals  <- raster::extract(rastAbundMed[[b]],originSitesProj)
)

str(weekly_region_vals)

sumByWeek <- lapply(weekly_region_vals, 
                    FUN = function(x){apply(x, 2, sum, na.rm = TRUE)})

regionSum <- lapply(sumByWeek, max)

relAbundYEWABreed <- prop.table(unlist(regionSum))

```


We now have all the data in order to run the `estTransition` or `estStrength` 
functions. 

Note - there was no overlap of data sources for individuals, i.e., we only had 
one type of data for each bird - so we set the overlap setting to “none”.

```{r}
## Run analysis for psi
system.time((psiYEWA <- estTransition(originSites = YEWA_origin_sites,
                                     targetSites = YEWA_target_sites,
                                     originPoints = originPointsYEWA,
                                     targetPoints = targetPointsYEWA,
                                     originAssignment = YEWA_genetics,
                                     originNames = originNames,
                                     targetNames = targetNames,
                                     nSamples = 10,
                                     isGL = isGLYEWA,
                                     isTelemetry = isTelemetryYEWA,
                                     isRaster = isRasterYEWA,
                                     isProb = isProbYEWA,
                                     captured = capturedYEWA,
                                     geoBias = geoBiasYEWA,
                                     geoVCov = geoVCovYEWA,
                                     originRaster = isoRasterYEWA,
                                     verbose = 2, 
                                     maxTries = 400,
                                     resampleProjection = st_crs(YEWA_origin_sites),
                                     nSim = 10,
                                     dataOverlapSetting = "none",
                                     targetRelAbund = relAbundYEWAWint)))



# Take a look at the results # 
psiYEWA
```

```{r echo = FALSE}

yewa_data <- list(originSites = YEWA_origin_sites,
                                     targetSites = YEWA_target_sites,
                                     originPoints = originPointsYEWA,
                                     targetPoints = targetPointsYEWA,
                                     originAssignment = YEWA_genetics,
                                     originNames = originNames,
                                     targetNames = targetNames,
                                     nSamples = 10,
                                     isGL = isGLYEWA,
                                     isTelemetry = isTelemetryYEWA,
                                     isRaster = isRasterYEWA,
                                     isProb = isProbYEWA,
                                     captured = capturedYEWA,
                                     geoBias = geoBiasYEWA,
                                     geoVCov = geoVCovYEWA,
                                     originRaster = isoRasterYEWA,
                                     verbose = 2, 
                                     maxTries = 400,
                                     resampleProjection = st_crs(YEWA_origin_sites),
                                     nSim = 10,
                                     dataOverlapSetting = "none",
                                     targetRelAbund = relAbundYEWAWint,
                                     originRelAbund = relAbundYEWABreed)

# saveRDS(yewa_data,"yewa_estTrans_data.rds")
```

Plot the results using the built-in plotting function
```{r, fig.width=5, fig.height=5, fig.cap="Estimated transition probabilities of the Yellow Warbler between breeding and nonbreeding periods"}
par(mar = c(2,3,0,0))
plot(psiYEWA, legend = "topleft", cex = 0.4)
```


## Calculating location bias and covariance structure to resample light-level geolocator data 

#### SGAT {#SGAT_bias}

The code below is a repeatable example for how to calculate location bias and location error using coordinates derived from light-level geolocators (see @cohen_quantifying_2018) analyzed using the
SGAT package [@wotherspoon_sgat_2017; @lisovski_lightlevel_2020]. 
    
```{r, eval = FALSE}
# Load in projections
data("projections")

# Define deployment locations (winter) # 
captureLocations<-matrix(c(-77.93,18.04,   # Jamaica
                           -80.94,25.13,   # Florida
                           -66.86,17.97),  # Puerto Rico
                            nrow=3, ncol=2, byrow = TRUE)

colnames(captureLocations) <- c("Longitude","Latitude")

# Convert capture locations into SpatialPoints #

CapLocs <- sf::st_as_sf(data.frame(captureLocations),
                      coords = c("Longitude","Latitude"),
                      crs = 4326)

# Project Capture locations # 

CapLocsM<-sf::st_transform(CapLocs, 'ESRI:54027')

# Retrieve raw non-breeding locations from github 
# First grab the identity of the bird so we can loop through the files 
# For this example we are only interested in the error 
# around non-breeding locations 
# here we grab only the birds captured during the non-breeding season 
# Using paste0 for vignette formatting purposes

winterBirds <- dget(paste0("https://raw.githubusercontent.com/",
                    "SMBC-NZP/MigConnectivity/master/",
                    "data-raw/GL_NonBreedingFiles/winterBirds.txt"))

# create empty list to store the location data #
Non_breeding_files <- vector('list',length(winterBirds))

# Get raw location data from Github #
for(i in 1:length(winterBirds)){
Non_breeding_files[[i]] <- dget(paste0("https://raw.githubusercontent.com/",
                                        "SMBC-NZP/MigConnectivity/master/data-raw/",
                                        "GL_NonBreedingFiles/NonBreeding_",
                                        winterBirds[i],".txt"))
}

# Remove locations around spring Equinox and potential migration points
# same NB time frame as Hallworth et al. 2015 

# two steps because subset on shapefile doesn't like it in a single step

Non_breeding_files <- lapply(Non_breeding_files,
                      FUN = function(x){
                      month <- as.numeric(format(x$Date,format = "%m"))
                               x[which(month != 3 & month != 4),]})

   
Jam <- c(1:9)   # locations w/in list of winterBirds captured in Jamaica
Fla <- c(10:12) # locations w/in list of winterBirds in Florida
PR <- c(13:16)  # locations w/in list of winterBirds in Puerto Rico

# Turn the locations into shapefiles #

NB_GL <- lapply(Non_breeding_files, 
                FUN = function(x){
                  sf::st_as_sf(data.frame(x),
                               coords = c("Longitude",
                                          "Latitude"),
                               crs = 4326)})

# Project into UTM projection #

NB_GLmeters <- lapply(NB_GL,
                      FUN = function(x){sf::st_transform(x,'ESRI:54027')})

# Process to determine geolocator bias and variance-covariance in meters #

# generate empty vector to store data #
LongError<-rep(NA,length(winterBirds)) 
LatError<-rep(NA,length(winterBirds))  

# Calculate the error in longitude derived 
# from geolocators from the true capture location 

LongError[Jam] <- unlist(lapply(NB_GLmeters[Jam],
                         FUN = function(x){mean(sf::st_coordinates(x)[,1]-
                                                  sf::st_coordinates(CapLocsM)[1,1])}))

LongError[Fla] <- unlist(lapply(NB_GLmeters[Fla],
                         FUN = function(x){mean(sf::st_coordinates(x)[,1]-
                                                  sf::st_coordinates(CapLocsM)[2,1])}))

LongError[PR] <- unlist(lapply(NB_GLmeters[PR],
                        FUN = function(x){mean(sf::st_coordinates(x)[,1]-
                                                 sf::st_coordinates(CapLocsM)[3,1])}))

# Calculate the error in latitude derived from
# geolocators from the true capture location 

LatError[Jam] <- unlist(lapply(NB_GLmeters[Jam],
                        FUN = function(x){mean(sf::st_coordinates(x)[,2]-
                                                  sf::st_coordinates(CapLocsM)[1,2])}))

LatError[Fla] <- unlist(lapply(NB_GLmeters[Fla],
                        FUN = function(x){mean(sf::st_coordinates(x)[,2]-
                                                 sf::st_coordinates(CapLocsM)[2,2])}))

LatError[PR] <- unlist(lapply(NB_GLmeters[PR],
                        FUN = function(x){mean(sf::st_coordinates(x)[,2]-
                                                  sf::st_coordinates(CapLocsM)[3,2])}))

# Get co-variance matrix for error of 
# known non-breeding deployment sites 

# lm does multivariate normal models if you give it a matrix dependent variable!

geo.error.model <- lm(cbind(LongError,LatError) ~ 1) 

geo.bias <- coef(geo.error.model)
geo.vcov <- vcov(geo.error.model)
``` 
  
#### FLightR {#FLightR_bias}

The code below is a repeatable example for how to calculate location bias and location error using location estimates derived from light-level geolocators analyzed in FLightR [@rakhimberdiev_flightr_2017; @lisovski_lightlevel_2020]. The geolocator data were processed and reported by @witynski_crosswise_2018.

```{r, eval = FALSE}
# Capture locations #
# See dates and summary data from Table 1. page 44.
# Maine #129
# Maine #132
# Maine #136
# Maine #137
# Wisconsin #139
# Wisconsin #144
# Wisconsin #146

caplocs <- cbind(c(rep(-70.6142,4),rep(-87.8123,3)),
                 c(rep(42.9891,4),rep(42.5000,3)))

# Departures #
departs <- as.POSIXct(c("2015-09-04","2015-08-22","2015-09-04","2015-08-28","2015-08-24","2015-08-24","2015-08-30"))

# read in the daily location data #
EstLocs <- read.csv("Yellow Warbler_Witynski (1).csv")

YEWAsf <- st_as_sf(EstLocs, coords = c("location.long","location.lat"), crs = 4326)

caplocs_sf <- st_as_sf(data.frame(bird = c(129,132,136,137,139,144,146),
                                  longitude = caplocs[,1],
                                  latitude = caplocs[,2]),
                       coords = c("longitude","latitude"),
                       crs = 4326)

# project into meters #
YEWAmeters <- sf::st_transform(YEWAsf,'ESRI:102010')
CapMeters <- sf::st_transform(caplocs_sf, 'ESRI:102010')

# split birds into their own data #

YEWAests <- split(YEWAmeters, f=YEWAmeters$'individual.local.identifier')

# Empirical estimates of bias and vcov, but on mean locations by bird
ests2 <- mapply(x = YEWAests,
                lats = st_coordinates(CapMeters)[,2],
                lng = st_coordinates(CapMeters)[,1],
                z = departs,
                FUN = function(x,lats,lng,z){
                  # keep only data from deployment until departure from breeding #
                  newdata <- x[x$timestamp<z,]

                  # calculate the difference in meters from estimated location
                  # to actual capture location
                  long <- mean(st_coordinates(newdata)[,1]) - lng
                  lat <- mean(st_coordinates(newdata)[,2]) - lats

                  return(c(difflong = long,
                                    difflat = lat))},

               SIMPLIFY = TRUE)

(geoBiasYEWA <- apply(ests2, 1, mean)) #Bias
#  difflong   difflat
# -6512.898 50964.969

apply(ests2, 1, var) #Variances
apply(ests2, 1, sd)

(geoVCovYEWA <- cov(t(ests2))) #Variance-covariance matrix

#            difflong    difflat
# difflong  259401516 -894387951
# difflat  -894387951 4708578511

```


### Literature Cited

<div id="refs"></div>
